---
title: "Penguins"
subtitle: "Statistics & Hypothesis testing"
output:
  html_document:
    keep_md: false
    df_print: paged
    css: www/style.css
    toc: true
    toc_float: 
          collapsed: false
          smooth_scroll: false
    number_sections: no
---

```{r message=FALSE, warning=FALSE, include=FALSE}

# get libraries / packages ---------------------------------------------------

library(palmerpenguins)  # install.packages("palmerpenguins")

library(tidyverse)
library(skimr)
library(infer)
library(e1071)  # to manage skewness

library(ggh4x)  #handle facet_wrap plots(https://teunbrand.github.io/ggh4x/articles/Facets.html) 
library(GGally) #correlaction plot

#for tables
library(knitr)
library(kableExtra)
library(bsicons)

# to manage HTML files
library(IRdisplay)
library(htmltools)


# call dataset ---------------------------------------------------

penguins <- palmerpenguins::penguins %>% 
  janitor::clean_names()

```

```{r message=FALSE, warning=FALSE, include=FALSE}
title_and_header_dir <- "www/title_and_header.html"


boxplot_info_dir <- "www/boxplot_info.html"

summary_outcomes_dir <- "www/summary_outcomes.html"
summary_skewness_outcomes_dir <- "www/summary_skewness_outcomes.html"
statistic_info_dir <- "www/summary_statistics_info.html"

hypothesis_test_outcome_dir <- "www/hypothesis_test_outcome.html"
hypothesis_test_info_dir <- "www/hypothesis_test_info.html"

correlation_outcomes_dir <- "www/correlation_outcomes.html"
correlation_info_dir <- "www/correlation_info.html"
```

```{r warning=FALSE, echo=FALSE, message=FALSE}

# Title & Header -----------------------------------------------------------

title_and_header <- readLines(title_and_header_dir, warn = FALSE)
HTML(paste(title_and_header, collapse = "\n"))

```

## 1. Exploratory Analysis

### 1.1. Variables

```{r warning=FALSE, echo=FALSE, message=FALSE}
variables_info <- data.frame(
  sapply(penguins,class)
)

variables_info$variable <- rownames(variables_info)
rownames(variables_info) <- 1:nrow(variables_info)
variables_info <- relocate(variables_info, variable)
variables_info$type <- variables_info$sapply.penguins..class.


variables_info <- variables_info %>% 
  select(variable, type) %>% 
  arrange(type) %>% 
  pivot_wider(names_from = variable, values_from = type)

variables_info %>% 
  kable(align = "c") %>% 
  kable_styling(
    full_width = TRUE,
    position = "center",
    bootstrap_options = c("none", "responsive")) %>% 
  row_spec(0, background = "#F5F2EB", color = "black") %>%
  row_spec(1:nrow(variables_info), extra_css = "border-bottom: 1px solid black")

```

### 1.2. Summary of Bill Lengths

```{r warning=FALSE, echo=FALSE, message=FALSE, out.width="100%"}

penguins %>% 
  group_by(species) %>% 
  skim(bill_length_mm) %>% 
  yank("numeric")

```

### 1.3. Scatter of Species Bill Lengths

```{r warning=FALSE, echo=FALSE, message=FALSE, out.width="100%"}

source(here::here("scripts/boxplot_analysis/boxplot_bill_length_distribution.R"), local = TRUE)$value

```



## 2. Summary statistics Analysis

```{r warning=FALSE, echo=FALSE, message=FALSE}

#call the function created to classify the skewness magnitude
source(here::here("scripts/function_magnitude_skewness_classification/magnitude_skewness_classification.R"), local = TRUE)

summary_statistics <- penguins %>% 
  drop_na() %>% 
  group_by(species) %>% 
  summarise(mean_bill_length     = round(mean(bill_length_mm), digits = 3),
            median_bill_length   = round(median(bill_length_mm), digits = 3),
            sd_bill_length       = round(sd(bill_length_mm), digits = 3),
            skewness_bill_length = round(skewness(bill_length_mm, 
                                            type = 1), digits = 3),
            magnitude_skewness_classification = 
              magnitude_skewness_classification(skewness_bill_length)
  )

summary_statistics %>% 
  kable(
    align = "c",
    col.names = c("Species", "Mean", "Median", "sd", "Skewness", "Skewness Classification")
  ) %>% 
  kable_styling(
    full_width = TRUE,
    position = "center",
    bootstrap_options = c("basic", "responsive")) %>% 
  # add_footnote(notation = "none", label = "Penguins Bill Length Summary Statistics") %>% 
  row_spec(0, background = "", color = "black", extra_css = "border-bottom:0px solid red") 
  # row_spec(1:3, extra_css = "border-bottom:1px solid black")


```


### 2.1. Mean & Median
The mean is more sensitive to outliers than the median. Upon examining the results, we observe a relatively small difference between the mean and median values for each species. This suggests either a limited presence of outliers or their absence altogether.

### 2.2. Standard Deviation (sd)
```{r warning=FALSE, echo=FALSE, message=FALSE}
summary_outcomes <- readLines(summary_outcomes_dir, warn = FALSE)
HTML(paste(summary_outcomes, collapse = "\n"))
```


### 2.3. Skewness
```{r warning=FALSE, echo=FALSE, message=FALSE}
summary_skewness_outcomes <- readLines(summary_skewness_outcomes_dir, warn = FALSE)
HTML(paste(summary_skewness_outcomes, collapse = "\n"))

# source(here::here("scripts/skewness_analysis.R"), local = TRUE)$value
source(here::here("scripts/skewness_analysis.R"), local = TRUE)

gridExtra::grid.arrange(adelie, chinstrap, gentoo, ncol = 3)
```


## 3. Hypothesis test & Probability

A hypothesis test to check whether the mean bill length of Gentoo is significantly longer than the mean bill length of Adelie.

**H0:** The average of bill length of Gentoo specie is the same of the average of Adelie specie. $$H_0: \mu_{average \ gentoo \ bill \ length} - \mu_{average \ adelie \ bill \ length} = 0$$ **H1:** The average of bill length of Gentoo specie is longer than the average of Adeleie specie. $$H_1: \mu_{average \ gentoo \ bill \ length} -\mu_{average \ adelie \ bill \ length} > 0$$

```{r message=FALSE, warning=FALSE, include=FALSE}
hypothesis_test <- source(file = here::here("scripts/hypothesis_testing/hypothesis_test.R"), local = TRUE)
```

```{r warning=FALSE, echo=FALSE, message=FALSE, out.width="100%"}

summary_hypothesis_test <- as.data.frame(
  c(
    "observed statistic" = round(obs_statistic, digits = 3),
    "p-value" = p_value,
    "min" = round(min(null_distribution$stat), digits = 3),
    "max"  = round(max(null_distribution$stat), digits = 3)
  )
)

summary_hypothesis_test %>% 
  kable(
    col.names = c("Observed Statistic", "p-value", "min", "max"),
    align = c("c")
  ) %>% 
  kable_styling(
    full_width = TRUE,
    position = "center",
    bootstrap_options = c("none", "responsive")) %>% 
  add_footnote(threeparttable = TRUE,
               # label = c("Results of Hypothesis Testing"), 
               escape = FALSE,
               notation = "none") %>% 
  row_spec(0, background = "#F5F2EB", color = "black") %>%
  row_spec(1, extra_css = "border-bottom: 1px solid black")

```

```{r warning=FALSE, echo=FALSE, message=FALSE}
hypothesis_test_outcome <- readLines(hypothesis_test_outcome_dir, warn = FALSE)
HTML(paste(hypothesis_test_outcome, collapse = "\n"))
```

```{r  warning=FALSE, echo=FALSE, message=FALSE, out.width="100%"}
hypothesis_test$value
```

## 4. Correlation Analysis

### 4.1. Correlation Matrix

```{r warning=FALSE, echo=FALSE, message=FALSE}
correlation_outcomes <- readLines(correlation_outcomes_dir, warn = FALSE)
HTML(paste(correlation_outcomes, collapse = "\n"))
```

```{r warning=FALSE, echo=FALSE, message=FALSE, out.height="100%"}

source(here::here("scripts/correlation_analysis/correlation_matrix.R"), local = TRUE)$value

```

### 4.2. Correlation

```{r  warning=FALSE, echo=FALSE, message=FALSE, out.width="100%"}
source(here::here("scripts/correlation_analysis/correlation_analysis_body.R"), local = TRUE)$value
```

```{r warning=FALSE, echo=FALSE, message=FALSE, out.width="100%"}
source(here::here("scripts/correlation_analysis/correlation_analysis_by_specie.R"), local = TRUE)$value
```

## 5. Auxiliar - Theory concepts
Expand each option below to get more details about the theory concepts that support the analysis. 
```{r warning=FALSE, echo=FALSE, message=FALSE}

boxplot_info <- readLines(boxplot_info_dir, warn = FALSE)
HTML(paste(boxplot_info, collapse = "\n"))

```

```{r warning=FALSE, echo=FALSE, message=FALSE}

statistic_info <- readLines(statistic_info_dir, warn = FALSE)
HTML(paste(statistic_info, collapse = "\n"))

```

```{r warning=FALSE, echo=FALSE, message=FALSE}

hypothesis_test_info <- readLines(hypothesis_test_info_dir, warn = FALSE)
HTML(paste(hypothesis_test_info, collapse = "\n"))

```

```{r warning=FALSE, echo=FALSE, message=FALSE}

correlation_info <- readLines(correlation_info_dir, warn = FALSE)
HTML(paste(correlation_info, collapse = "\n"))

```
